A new simplified manual tour, with examples in mathematica

### 2022-04-01 paper/paper.Rmd
started intro

%\doublespacing
# Introduction

<!-- Background -->

>>>>

<!-- \doublespacing-->
# Introduction

<!-- Background -->
From a statistical perspective it is rare to have data that are strictly 3D, and so unlike in most computer graphics applications, the more useful methods for data analysis show projections from an arbitrary dimensional space. These are dynamic data visualizations methods and are collected under the term *tours*. Tours involve views of high-dimensional ($p$) data in low-dimensional ($d$) projections. In his original paper on the grand tour, @As85 provided several algorithms for tour paths that could theoretically show the viewer the data *from all sides*. Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9. PRIM-9 had user-controlled rotations on coordinate axes, allowing one to manually tour through low-dimensional projections. It is impractical to impossible to steer through all possible projections, unlike Asimov's tours which allows one to quickly see many, many different projections. After Asimov there have been many, many tour developments, as summarized in @lee2021.

<!-- Manual tours -->
One such direction of work develops the ideas from PRIM-9, to provide manual control of a tour. @cook_manual_1997 describes controls for 1D (or 2D) projections, in a 2D (or 3D) manipulation space, allowing the user to select any variable axis, and rotate it into or out of or around the projection through horizontal, vertical, oblique, radial or angular changes in value. @spyrison_spinifex_2020 refines this algorithm and implements them to generate animations.

<!-- Purpose -->
Manual controls are especially useful for assessing sensitivity of structure to particular elements of the projection. There are many places where it is useful. In exploratory data analysis, where one sees clusters in a projection, can some variables be removed from the projection without affecting the clustering. For interpreting models, one can reduce or increase a variable's contribution to examine the variable importance. These controls can also be used to interactively generate facetted plots [@XXX], or spatiotemporal glyphmaps [@XXX].

### 2022-04-01 paper/paper.tex
started intro

\%\doublespacing \# Introduction

@@

>>>>

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

From a statistical perspective it is rare to have data that are strictly
3D, and so unlike in most computer graphics applications, the more
useful methods for data analysis show projections from an arbitrary
dimensional space. These are dynamic data visualizations methods and are
collected under the term \emph{tours}. Tours involve views of
high-dimensional (\(p\)) data in low-dimensional (\(d\)) projections. In
his original paper on the grand tour, \citet{As85} provided several
algorithms for tour paths that could theoretically show the viewer the
data \emph{from all sides}. Prior to Asimov's work, there were numerous
preparatory developments including \citet{tukey}'s PRIM-9. PRIM-9 had
user-controlled rotations on coordinate axes, allowing one to manually
tour through low-dimensional projections. It is impractical to
impossible to steer through all possible projections, unlike Asimov's
tours which allows one to quickly see many, many different projections.
After Asimov there have been many, many tour developments, as summarized
in \citet{lee2021}.

One such direction of work develops the ideas from PRIM-9, to provide
manual control of a tour. \citet{cook_manual_1997} describes controls
for 1D (or 2D) projections, in a 2D (or 3D) manipulation space, allowing
the user to select any variable axis, and rotate it into or out of or
around the projection through horizontal, vertical, oblique, radial or
angular changes in value. \citet{spyrison_spinifex_2020} refines this
algorithm and implements them to generate animations.

Manual controls are especially useful for assessing sensitivity of
structure to particular elements of the projection. There are many
places where it is useful. In exploratory data analysis, where one sees
clusters in a projection, can some variables be removed from the
projection without affecting the clustering. For interpreting models,
one can reduce or increase a variable's contribution to examine the
variable importance. These controls can also be used to interactively
generate facetted plots \citep{XXX}, or spatiotemporal glyphmaps
\citep{XXX}.

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/manualsequence-1.pdf}
\caption{Sequence of projections where V3 contribution is changed.}
\end{figure}
